Counterfactual Training: Teaching Models Plausible and Actionable Explanations

### 2025-03-06 paper/quarto_ecml/paper.tex
long road ahead still

\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage[misc]{ifsym}
\newcommand{\corr}{(\Letter)}
\usepackage{placeins}
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Fig.}
\else
  \newcommand\figurename{Fig.}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
% \usepackage{amsthm}
% \theoremstyle{plain}
% \newtheorem{corollary}{Research Question}[section]
% \theoremstyle{plain}
% \newtheorem{proposition}{Proposition}[section]
% \theoremstyle{definition}
% \newtheorem{example}{Example}[section]
% \theoremstyle{definition}
% \newtheorem{definition}{Definition}[section]
% \theoremstyle{remark}
% \AtBeginDocument{\renewcommand*{\proofname}{Proof}}
% \newtheorem*{remark}{Remark}
% \newtheorem*{solution}{Solution}
% \newtheorem{refremark}{Remark}[section]
% \newtheorem{refsolution}{Solution}[section]
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\newcounter{quartocalloutnteno}
\newcommand{\quartocalloutnte}[1]{\refstepcounter{quartocalloutnteno}\label{#1}}

@@

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Counterfactual Training: Teaching Models Plausible and Actionable Explanations},
  pdfauthor={Patrick Altmeyer; Aleksander Buszydlik; Arie van Deursen; Cynthia C. S. Liem},
  pdfkeywords={Counterfactual Explanations, Explainable
AI, Representation Learning},
  colorlinks=true,
  linkcolor={black},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\keywords{Counterfactual Explanations \and Explainable
AI \and Representation Learning}

@@

@@

\begin{tcolorbox}[enhanced jigsaw, titlerule=0mm, bottomrule=.15mm, coltitle=black, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, opacityback=0, bottomtitle=1mm, leftrule=.75mm, breakable, colframe=quarto-callout-note-color-frame, left=2mm, colback=white, toptitle=1mm, toprule=.15mm, rightrule=.15mm, opacitybacktitle=0.6, title={Note}]

@@

\begin{tcolorbox}[enhanced jigsaw, titlerule=0mm, bottomrule=.15mm, coltitle=black, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, opacityback=0, bottomtitle=1mm, leftrule=.75mm, breakable, colframe=quarto-callout-note-color-frame, left=2mm, colback=white, toptitle=1mm, toprule=.15mm, rightrule=.15mm, opacitybacktitle=0.6, title={Note \ref*{nte-gen-params-final-run-train}: Training Phase}]

@@

\begin{tcolorbox}[enhanced jigsaw, titlerule=0mm, bottomrule=.15mm, coltitle=black, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, opacityback=0, bottomtitle=1mm, leftrule=.75mm, breakable, colframe=quarto-callout-note-color-frame, left=2mm, colback=white, toptitle=1mm, toprule=.15mm, rightrule=.15mm, opacitybacktitle=0.6, title={Note \ref*{nte-gen-params-final-run-eval}: Evaluation Phase}]

>>>>

\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage[misc]{ifsym}
\newcommand{\corr}{(\Letter)}
\usepackage{placeins}
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Fig.}
\else
  \newcommand\figurename{Fig.}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
% \usepackage{amsthm}
% \theoremstyle{definition}
% \newtheorem{example}{Example}[section]
% \theoremstyle{plain}
% \newtheorem{proposition}{Proposition}[section]
% \theoremstyle{plain}
% \newtheorem{corollary}{Research Question}[section]
% \theoremstyle{definition}
% \newtheorem{definition}{Definition}[section]
% \theoremstyle{remark}
% \AtBeginDocument{\renewcommand*{\proofname}{Proof}}
% \newtheorem*{remark}{Remark}
% \newtheorem*{solution}{Solution}
% \newtheorem{refremark}{Remark}[section]
% \newtheorem{refsolution}{Solution}[section]
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\newcounter{quartocalloutnteno}
\newcommand{\quartocalloutnte}[1]{\refstepcounter{quartocalloutnteno}\label{#1}}

@@

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Counterfactual Training: Teaching Models Plausible and Actionable Explanations},
  pdfauthor={Patrick Altmeyer; Aleksander Buszydlik; Arie van Deursen; Cynthia C. S. Liem},
  pdfkeywords={Counterfactual Explanations, Algorithmic
Recourse, Explainable AI, Adversarial Robustness, Representation
Learning},
  colorlinks=true,
  linkcolor={black},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\keywords{Counterfactual Explanations \and Algorithmic
Recourse \and Explainable AI \and Adversarial
Robustness \and Representation Learning}

@@

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

@@

\begin{tcolorbox}[enhanced jigsaw, left=2mm, coltitle=black, colbacktitle=quarto-callout-note-color!10!white, rightrule=.15mm, leftrule=.75mm, arc=.35mm, titlerule=0mm, opacitybacktitle=0.6, toptitle=1mm, colback=white, title={Note}, bottomtitle=1mm, bottomrule=.15mm, toprule=.15mm, opacityback=0, colframe=quarto-callout-note-color-frame, breakable]

@@

\begin{tcolorbox}[enhanced jigsaw, left=2mm, coltitle=black, colbacktitle=quarto-callout-note-color!10!white, rightrule=.15mm, leftrule=.75mm, arc=.35mm, titlerule=0mm, opacitybacktitle=0.6, toptitle=1mm, colback=white, title={Note \ref*{nte-gen-params-final-run-train}: Training Phase}, bottomtitle=1mm, bottomrule=.15mm, toprule=.15mm, opacityback=0, colframe=quarto-callout-note-color-frame, breakable]

@@

\begin{tcolorbox}[enhanced jigsaw, left=2mm, coltitle=black, colbacktitle=quarto-callout-note-color!10!white, rightrule=.15mm, leftrule=.75mm, arc=.35mm, titlerule=0mm, opacitybacktitle=0.6, toptitle=1mm, colback=white, title={Note \ref*{nte-gen-params-final-run-eval}: Evaluation Phase}, bottomtitle=1mm, bottomrule=.15mm, toprule=.15mm, opacityback=0, colframe=quarto-callout-note-color-frame, breakable]
